---
title:  "Estimation of changes in growth"
subtitle: "Using a state-space growth model"
author: "Christine Stawitz<br>NOAA Fisheries, OST<br>Seattle, WA<br>`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides-style.css", "tufte.css"]
    lib_dir: libs
    nature:
      ratio: '4:3'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc( (%current% - 1) / (%total% - 1) * 100%);">
          </div>
        </div>`
---
layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
require(here)
require(dplyr)
require(tidyr)
require(ggplot2)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
xaringanExtra::use_tile_view()
```

---

# Introduction
* Different papers have come to different conclusions regarding growth change
* Regional differences?
* Species differences? 
* Modeling differences?

---
# ICES/PICES Working Group on Growth Change 

California current species modeled using multiple frameworks
Acknowledge members

---
# State-space autoregressive length-at-age model
* Stawitz et al. 2014
* Adapted for the Stan software by Sean Anderson and Christine Stawitz
* Added ability to estimate different variation processes simultaneously

---

# The California Current

---
# Estimation settings
* Needed a strong prior on $\sigma_p$ (process error) and $\sigma_o$ (observation error) since the data can't distinguish between the two
* 4 chains
* 1000 warmup iterations

---
# Model types
* Initial size
* Cohort
* Annual


---
# Results

How many stocks had significant effects ow which type?

---
# Results

```{r show_example, echo = FALSE}
knitr::include_graphics("../plots/sablefish/all-effects.pdf")

```


Sablefish - giving $\sigma_p$ very narrow prior on 0.2 fixed BFMI warnings
- removed initial size effects since these were very small
Lingcod, even with fixed 0.2 had EBFMI less than 0.2
---
# Results
---
# Next steps
---